---
title: "Predictive Modeling - Popular YouTube Videos"
author: "Alfred Prah"
output: html_notebook
---

# Trending YouTube Video Statistics

YouTube (the world-famous video sharing website) maintains a list of the top trending videos on the platform. My dataset includes the YouTube video title, channel title, publish time, tags, views, likes and dislikes, description, and comment count. The goal is to use the available data about trending videos to determine which video uploads will trend in the United States of America. 

## Definitions of variables:

video_id - this is pre-determined by YouTube. For the purposes of my predictive model, it is not a parameter to worry about. 

trending_date - This refers to the date a particular YouTube video was "trending" 

title - This refers to the title assigned to a YouTube video by the people uploading it to their channel. 

channel_title - This refers to the name of the channel the YouTube video is being uploaded through. 

category_id - This refers to the category that a YouTube video falls into. 

publish_time - This refers to the time that the YouTube video was published on the channel. 

tags - Tags are words and phrases used to give YouTube viewers and algorithms context about a video. 

views - This refers to the number of people who watch a particular video. 

likes - This refers to the number of people who vote a video up. Ie, they click on the "thumbs up" icon underneath the video. 

dislikes - This refers to the number of people who vote a video down. Ie, they click on the "thumbs down" icon underneath the video.

comment_count - This refers to how many comments a particular video received from its viewers. 

thumbnail_link - This refers to the picture that was used a "cover photo" for the uploaded video. 

comments_disabled - This column provides an account of videos that had the comment function disabled. 

ratings_disabled - This refers to whether a channel gives its viewers the opportunity to vote a video up or down.

video_error_or_removed - This refers to whether the formerly known link to a video has been changed/removed, or whether there is any error in accessing the video content. 

description - This refers to video descriptions that channel owners use to preview their video to prospective viewers. 

## Predictive Modeling Problem
I would like to predict popular YouTube videos.


## libraries
```{r libraries, warning=FALSE, message=FALSE}
library(readr)
library(lubridate)
library(tidyverse)
library(assertr)
library(janitor)
library(chron)
```

## Data Import 
```{r import the data set, warning=FALSE, message=FALSE}
video_statistics <- read_csv("youtube-new/USvideos.csv")
```

## Data Inspection
```{r data inspection, warning=FALSE, message=FALSE}
str(video_statistics)
dim(video_statistics)
video_statistics
```

## Data Testing
```{r data testing: check for NA values in all columns}
video_statistics %>%
  assert(not_na, everything())  # check whether there are NA values
```
assetr stopped execution because there are 578 NA values in the "description" column. This is good information and something I would take into account, going forward.


```{r data testing: check for "TRUE" or "FALSE in the "comments_disabled" column}
#check to see if the values of this column are either TRUE or FALSE, exclusively. 
assert(video_statistics, in_set("FALSE", "TRUE"), comments_disabled)  
```
assetr did not stop execution. This means that the "comments_disabled" column contains only True or False values. 

```{r data testing: check for "TRUE" or "FALSE in the "ratings_disabled" column}
#check to see if the values of this column are either TRUE or FALSE, exclusively.
assert(video_statistics, in_set("FALSE", "TRUE"), ratings_disabled)  
```
assetr did not stop execution. This means that the "ratings_disabled" column contains only True or False values.

```{r data testing: check for "TRUE" or "FALSE in the "video_error_or_removed" column}
##check to see if the values of this column are either TRUE or FALSE, exclusively.
assert(video_statistics, in_set("FALSE", "TRUE"), video_error_or_removed)  
```
assetr did not stop execution. This means that the "video_error_or_removed" column contains only True or False values.

## Data Cleaning
```{r data cleaning: change TRUE or FALSE values in "comments_disabled" to T or F}
video_statistics <- video_statistics %>%
  mutate(comments_disabled = substring(comments_disabled, 1, 1))
```

```{r data cleaning: change TRUE or FALSE values in "ratings_disabled" to T or F}
video_statistics <- video_statistics %>%
  mutate(ratings_disabled = substring(ratings_disabled, 1, 1))
```

```{r data cleaning: change TRUE or FALSE values in "video_error_or_removed" to T or F}
video_statistics <- video_statistics %>%
  mutate(video_error_or_removed = substring(video_error_or_removed, 1, 1))
```


## Feature Engineering 
I would like to assign a 1 or 0 to the True/False values of the "comments_disabled" column. This is because most statistical and machine learning models cannot make sense of string inputs, and one-hot encoding is one way to numerically represent contextual information.
```{r feature engineering: assign a 1 or 0 to values in "comments_disabled" column}
video_statistics <- video_statistics %>%
  mutate(if_comment_disabled = ifelse(comments_disabled == "T", 1, 0))

#function test
video_statistics %>% 
  assert(in_set(1,0), if_comment_disabled) #the column contains either a 1 or 0, exclusively.
```


I would like to assign a 1 or 0 to the True/False values of the "ratings_disabled" column. This is because most statistical and machine learning models cannot make sense of string inputs, and one-hot encoding is one way to numerically represent contextual information.
```{r feature engineering: assign a 1 or 0 to values in "ratings_disabled" column}
video_statistics <- video_statistics %>%
  mutate(if_ratings_disabled = ifelse(ratings_disabled == "T", 1, 0))

#function test
video_statistics %>% 
  assert(in_set(1,0), if_ratings_disabled) #the column contains either a 1 or 0, exclusively.
```

I would like to assign days of the week to see if the day of the week that the video was uploaded plays any role in determining how popular it will become.
```{r feature engineering: add a column for the day of the week that each video was}
video_statistics <- video_statistics %>%
  mutate(week_day = wday(publish_time)) 

#function test
video_statistics %>% 
  assert(within_bounds(1,7, publish_time)) # this confirms that the weekdays in my column have values 1 through 7.
```

Create a column to determine what time of day the videos are published on YouTube.
I would like to determine if there is a trend of the time of day that trending videos are published. Ie, are the popular videos generally published in the mornings, afternoons or evenings? 
```{r feature engineering: separate the time of upload into a new column for further analysis}
upload_time <-
  format(as.POSIXct(video_statistics$publish_time) , format = "%H:%M:%S")

video_statistics <- video_statistics %>%
  mutate(upload_time = upload_time) 


#function test

#the %>% operation doesn't seem to work here, oops!
assert(video_statistics, is.timepoint(video_statistics$publish_time)) #the column contains only "time" values.
```


```{r feature engineering: assign time of day to "upload_time", warning=FALSE, message=FALSE}
#create intervals for the various times of day: morning, afternoon, evening, night

interval1 <- interval(as_datetime(hms("5:00:00")), as_datetime(hms("11:59:59")))
interval2 <- interval(as_datetime(hms("12:00:00")), as_datetime(hms("16:59:59")))
interval3 <- interval(as_datetime(hms("17:00:00")), as_datetime(hms("20:59:59")))
interval4 <- interval(as_datetime(hms("21:00:00")), as_datetime(hms("23:59:59")))
interval5 <- interval(as_datetime(hms("00:00:00")), as_datetime(hms("4:59:59")))

#place the various times of day into 4 separate columns: morning, afternoon, evening, night 
video_statistics <- video_statistics %>%
  mutate(
    time_of_day =
      case_when(
        as_datetime(hms(upload_time)) %within% interval1 ~ "morning",
        as_datetime(hms(upload_time)) %within% interval2 ~ "afternoon",
        as_datetime(hms(upload_time)) %within% interval3 ~ "evening",
        as_datetime(hms(upload_time)) %within% interval4 ~ "night",
        as_datetime(hms(upload_time)) %within% interval5 ~ "night"
        
      )
  )
```



```{r feature engineering: create new columns for the time of day of each upload}
#assign a 1 to a table each table that corresponds to the time of day that a particular video was uploaded. 

video_statistics <- video_statistics %>%
  mutate(
    # Create morning column
    morning = ifelse(time_of_day == "morning", 1, 0),
    
    # Create afternoon column
    afternoon = ifelse(time_of_day == "afternoon", 1, 0),
    
    # Create evening column
    evening = ifelse(time_of_day == "evening", 1, 0),
    
    #create night column
    night = ifelse(time_of_day == "night", 1, 0)
  )

assert(video_statistics, in_set(1, 0), morning) #the new column contains either a 1 or 0, exclusively.
assert(video_statistics, in_set(1, 0), afternoon) #the new column contains either a 1 or 0, exclusively.
assert(video_statistics, in_set(1, 0), evening) #the new column contains either a 1 or 0, exclusively.
assert(video_statistics, in_set(1, 0), night) #the new column contains either a 1 or 0, exclusively.
```

## Modeling
Most statistical and machine learning models cannot make sense of string inputs, and one-hot encoding is one way to numerically represent contextual information. To prepare my data and its inputs for modeling, I have ensured that the the parameters of interest were converted to numerical values during the "feature-engineering" process. I have also re-arranged the variables in the dataset to place the ones of interest at the front. 
```{r modeling: re-arrange, with parameters of interest at the front}
video_statistics <- video_statistics %>%
  select(
    video_id,
    if_comment_disabled,
    if_ratings_disabled,
    week_day,
    morning,
    afternoon,
    evening,
    night,
    upload_time,
    time_of_day,
    everything()
  )
```