---
title: "assign4"
output: html_notebook
---

### These are the libraries I will be using:
```{r}
library(lubridate)
library(stringr)
library(RVerbalExpressions)
```


For the regular expression exercises, try writing first using regular expressions directly, then try using rverbalexpressions. Provide the code for both. 

# Part 1

## R4DS 16.4.5

Create a vector of dates giving the first day of every month in the current year. 
```{r the first day of every monthj}
floor_date(today(), unit = "year") + months(0:11)
```


Write a function that given your birthday (as a date), returns how old you are in years.
```{r function that returns current age}
my_age <- function(birthday) {
  (birthday %--% today()) %/% years(1)
}
my_age(ymd("1997-01-19"))
```

Write a function that given your birthday, returns the day of the week you were born on. 
```{r day of the week born}
day_born <- function(x){
 day_of_week <- wday(x,label = T)
  return(day_of_week)
}
day_born(mdy(01191997))
```


## R4DS 14.3.2.1

Given the corpus of common words in stringr::words, create regular expressions that find all words that have seven letters or more. (Since this list is long, you might want to use the match argument to str_view() to show only the matching or non-matching words.)
```{r words that have 7 letters or more}
str_view(stringr::words, ".......", match = TRUE)

seven_or_more <- rx() %>% 
  rx_start_of_line() %>% 
  rx_word_char() %>%
  rx_repeat_previous(7) %>%
  rx_anything() %>% 
  rx_end_of_line() %>% 
  print()
str_view(words,seven_or_more, match = T)
```
## R4DS 14.3.3.1

Use str_detect to find all words that:

That only contain consonants. (Hint: thinking about matching “not”-vowels.)
```{r words containing only consonants}
str_subset(stringr::words, "^[^aeiou]+$")

find_c <- rx() %>% 
  rx_start_of_line() %>% 
  rx_something_but("aeiou") %>% 
  rx_end_of_line()
find_c
words[str_detect(words, find_c)]
```


Empirically verify the rule “i before e except after c”.
```{r rule verification: “i before e except after c”}
str_subset(stringr::words, "(cei|[^c]ie)")
str_subset(stringr::words, "(cie|[^c]ei)")

rule_verify <- rx() %>% 
  rx_something_but("c") %>% 
  rx_find("i") %>% 
  rx_find("e") %>% 
  print()
words[str_detect(words, rule_verify)]
```

Create a regular expression that will match telephone numbers as commonly written in the US. Split this up into multiple expressions as needed.
```{rnregex to match telephone number}

phone_number <-  rx() %>% 
  rx_start_of_line() %>%
  rx_maybe("(") %>%
  rx_digit() %>%
  rx_repeat_previous(3) %>%
  rx_maybe(")") %>%
  rx_any_of(" .-") %>%
  rx_digit() %>%
  rx_repeat_previous(3) %>%
  rx_any_of(" .-") %>%
  rx_digit() %>%
  rx_repeat_previous(4) %>%
  rx_end_of_line() 
 
 grepl(phone_number, 815-585-6481)

```

# Part 2


Choose a dataset (this can be the same dataset as the last assignment, or a new one). Define a predictive modeling problem. Create the appropriate files and notebooks to do the following:

1. Describe the data and the problem

2. Read in and check data

2. Clean up the data. 

*Note: You may request that score for these sections replace your score for the coding portion of Assignment 3.*

4. Using best practices, write four functions which add engineered features to the dataset, including a description and rationale. Include a test to determine whether the features are working as intended. 

5. Prepare the data for modeling

Note: this will form the basis for your midterm project. Take advantage of TA's and my office hours. We can provide feedback and guidance. 

